---
title: "Approximate Confidence Intervals"
format: html
editor: visual
---

## Wald interval

In class, we decided to consider $X_1,\ldots,X_{38}\sim\operatorname{Bernoulli}(0.4)$.

Even though we secretly know $p = 0.4$, let's estimate it with a Wald interval.  In particular, if $\hat p$ is the sample proportion (i.e., the number of 1's we get divided by 38), then we want the following to have probability $1 - \alpha$.

$$\hat p - z_{\alpha/2}\sqrt{\frac{\hat p(1 - \hat p)}{n}} <  p < \hat p + z_{\alpha/2}\sqrt{\frac{\hat p(1 - \hat p)}{n}}$$

Let's test this with a Monte Carlo simulation.  Choosing $1 - \alpha = 0.95$, we compute a bunch of confidence intervals and expect that about 95\% of the time, we'll get one containing $p = 0.4$.


```{r}
z <- qnorm(0.05/2, lower.tail = FALSE)
n <- 38
p <- 0.4
mean(replicate(10000, {
    x <- rbinom(n, 1, p);
    pHat <- mean(x);
    E <- z * sqrt(pHat * (1 - pHat) / n);
    pHat - E < p && p < pHat + E}))
```

## Wilson interval

Recall that the Wald interval uses two different levels of approximation: the central limit theorem (so we can use probabilities from a standard normal distribution) and approximating the $p$'s that show up in the standard deviation of the Bernoulli distribution with $\hat p$'s.

If we instead solve the inequality $|\hat p - p| < z_{\alpha/2}\sqrt{\frac{p(1-p)}{n}}$ for $p$, so that our only level of approximation is the central limit theorem, then we get the Wilson interval, bounded by:

$$\frac{1}{1+z_{\alpha/2}^2/n}\left(\hat p + \frac{z_{\alpha/2}^2}{2n} \pm z_{\alpha/2}\sqrt{\frac{\hat p(1-\hat p)}{n} + \frac{z_{\alpha/2}^2}{4n^2}}\right)$$

Let's play the same game and perform a Monte Carlo simulation for the Wilson interval:

```{r}
mean(replicate(10000, {
    x <- rbinom(n, 1, p);
    pHat <- mean(x);
    center <- 1/(1 + z^2/n) * (pHat + z^2/(2*n));
    radius <- 1/(1 + z^2/n) * z * sqrt(pHat * (1 - pHat)/n + z^2/(4*n^2));
    center - radius < p && p < center + radius}))
```


